### 2014-06-08 presentation/VCPSlides.tex
Presentation start

\documentclass{beamer}

\definecolor{OSUdkbrn}{RGB}{104,80,64}
\definecolor{OSUmedbrn}{RGB}{176,96,16}
\definecolor{OSUltbrn}{RGB}{224,195,152}
% from: http://oregonstate.edu/brand/color-palette

\mode<presentation>
{
	\usetheme{CambridgeUS}
	\setbeamercovered{transparent}
	\setbeamertemplate{footline}[page number]{}
	\setbeamercolor{palette tertiary}{fg=white, bg=OSUdkbrn}
	\setbeamercolor{title}{fg=OSUdkbrn}
	\setbeamercolor{frametitle}{fg=OSUdkbrn}
}

\title{Variable Circular Plots: Station Placement and the Independence Assumption}
\subtitle{Master's Research Project}
\author{Matt Edwards\\ Advisor: Claudio Fuentes}
\date{June 11, 2014}

\begin{document}
\begin{frame}
	\titlepage
\end{frame}

%\begin{frame}
%	\frametitle{Outline}
%	\tableofcontents
%\end{frame}

% -------------------------------- INTRODUCTION --------------------------------------------------
\section{Introduction}

\begin{frame}{Distance Sampling}
\begin{itemize}
\item Population Density Estimation
\item[]
\item Used often in Ecological Sciences
\item[]
\item Began with roadside surveys in early 1900s
\item[]
\item Theory developed through mid-1900s

\end{itemize}
\end{frame}

\begin{frame}{Distance Sampling: History}
\begin{itemize}
\item Emlen 1971
\item Ramsey, Scott 1979, 1980 Papers
\item Burnham, Anderson, and Jeffrey L. Laake 1980 Monograph
\item Buckland et. al. 1993, 2001
\item Continued Development into 2000s
\begin{itemize}
\item Combined with Mark-Recapture Methods
\item Extended for use with underwater acoustics for estimates of krill populations 
\item Combined with camera traps to estimate populations
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Distance Sampling}

\end{frame}

\end{document}